A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-28 text/manuscript.Rmd
Combine fig1+2, 3+4, 7+8
Update figure legend text

**Figure 1: Molecular pathway profiles of tumor samples with one or more
mutations.** Each of 377 selected Reactome pathways (rows) is classified as
disrupted if one or more genes is mutated in the tumor sample (columns). Red
cells denote pathway disruption. Tumors types reflect standardized abbreviations
from the TCGA project.

@@

#### Figure 2
```{r, child='../code/figure2.Rmd'}
```

**Figure 2: Clustering of tumor samples** Different rotational perspectives of
the same MCA-based UMAP projection in three-dimensional space. Each dot
corresponds to a tumor sample. The same colors indicate the tumor's cluster
identity throughout this manuscript.

@@

**Figure 3: Pathway-based clustering independent of tissue-of-origin** 
A) Sankey plot showing correspondence between cancer type and cluster identity.
B) Projection of breast cancer subtypes onto the UMAP. C) Projection of BRCA1/2
somatic mutation onto the UMAP. D) Projection of tumor stage onto the UMAP,
regardless of cancer type. E) Projection of metastatic status onto the UMAP.
Abbrevs: M0 = non-metastatic tumors, M1 = metastatic tumors.

@@

#### Figure 4
```{r child = '../code/figure4.Rmd'}
```

**Figure 4: Somatic mutation frequencies for each cluster** 
Vertical axis shows log mutation count, horizontal axis is cluster identity.
Each dot represents an individual tumor sample, ranked lowest to highest by
mutation count. The median mutation count in each cluster is indicated by the
horizontal line.

>>>>

**Figure 1: Clustering the molecular pathway profiles of tumor samples with one or more
mutations.** A) Each of 377 selected Reactome pathways (rows) is classified as
disrupted if one or more genes is mutated in the tumor sample (columns). Red
cells denote pathway disruption. Tumors types reflect standardized abbreviations
from the TCGA project. B) Different rotational perspectives of
the same MCA-based UMAP projection in three-dimensional space. Each dot
corresponds to a tumor sample. The same colors indicate the tumor's cluster
identity throughout this manuscript.

@@

@@

**Figure 3: Pathway-based clustering independent of tissue-of-origin** 
A) Sankey plot showing correspondence between cancer type and cluster identity.
B) Projection of breast cancer subtypes onto the UMAP. C) Projection of BRCA1/2
somatic mutation onto the UMAP. D) Projection of tumor stage onto the UMAP,
regardless of cancer type. E) Projection of metastatic status onto the UMAP.
Abbrevs: M0 = non-metastatic tumors, M1 = metastatic tumors. F) Somatic mutation 
frequencies for each cluster. Vertical axis shows log mutation count, horizontal 
axis is cluster identity. Each dot represents an individual tumor sample, ranked 
lowest to highest by mutation count. The median mutation count in each cluster is 
indicated by the horizontal line.

@@

### 2020-08-28 text/survival.Rmd
Combine fig1+2, 3+4, 7+8
Update figure legend text

---
output:
  word_document: default
  html_document: default
---
If our clusters represent biological states
distinct from tissue of origin, they may have different prognoses within cancer
types or across all cancers. These analyses are limited by confounding factors
of age and stage at diagnosis, sex, ethnicity, and tissue-specific disease
progression. To explore these ideas, we used Bayesian inference to test models
of survival using
public longevity data from the CDC.
Our model estimates the effect of cancer type and cluster-specific cancer
effects independently, resulting in a cancer and cluster-specific estimate of
the aging rate multiplier, $k$.

@@

**Figure 7: Comparison of different cancers, simulated and actual results
results.** To illustrate the risk model, we compare survival of four simulated
cohorts. A) A group consisting of randomly selected 30 year old women (light
blue) vs. an equal sized group of 30 year old women with cancer, k = 2 (light
orange). To illustrate how the mix of ages alters the survival curve, we compare
a randomly selected group of people with ages uniformly distributed from 30 to
70 (dark blue) and a randomly selected group with uniformly distributed ages and
cancer k = 2 (dark orange). B,C,D,E) We compare a randomly selected group with
correct age distribution for the given cancer (turqoise) to the model predicted
survival for the given average k (blue) and the actual survival for the patients
in our dataset (orange). Differences with actual data are likely due to a
mixture of cluster identities which we exclude from these simulations for
simplicity.

@@

#### Figure 8
```{r fig.width=3, message=FALSE, warning=FALSE}
pcanc+mytheme
```

**Figure 8 Tissue specific effects** Effects are compared between cancer type by
showing the probability density for the posterior value of the $k_{tis}$
multiplier. Larger values correspond to decreased age-independent life
expectancy.

>>>>

If our clusters represent biological states
distinct from tissue of origin, they may have different prognoses within cancer
types or across all cancers. These analyses are limited by confounding factors
of age and stage at diagnosis, sex, ethnicity, and tissue-specific disease
progression. To explore these ideas, we used Bayesian inference to test models
of survival using
public longevity data from the CDC.
Our model estimates the effect of cancer type and cluster-specific cancer
effects independently, resulting in a cancer and cluster-specific estimate of
the aging rate multiplier, $k$.

@@

**Figure 7: Comparison of different cancers, simulated and actual results
results.** To illustrate the risk model, we compare survival of four simulated
cohorts. A) A group consisting of randomly selected 30 year old women (light
blue) vs. an equal sized group of 30 year old women with cancer, k = 2 (light
orange). To illustrate how the mix of ages alters the survival curve, we compare
a randomly selected group of people with ages uniformly distributed from 30 to
70 (dark blue) and a randomly selected group with uniformly distributed ages and
cancer k = 2 (dark orange). B,C,D,E) We compare a randomly selected group with
correct age distribution for the given cancer (turqoise) to the model predicted
survival for the given average k (blue) and the actual survival for the patients
in our dataset (orange). Differences with actual data are likely due to a
mixture of cluster identities which we exclude from these simulations for
simplicity. F) Effects are compared between cancer type by
showing the probability density for the posterior value of the $k_{tis}$
multiplier. Larger values correspond to decreased age-independent life
expectancy.

@@
